---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
fontsize: 12pt
geometry: margin=1in

---

Do we need detailed demographic data to forecast population responses to climate change?
========================================================================================

### Andrew T. Tredennick and Peter B. Adler

*Andrew T. Tredennick ([atredenn@gmail.com](mailto:atredenn@gmail.com)), Department of Wildland Resources and the Ecology Center, Utah State University, Logan, UT*

*Peter B. Adler, Department of Wildland Resources and the Ecology Center, Utah State University, Logan, UT*

```{r libraris, include=FALSE}
library(ggplot2)
library(plyr)
library(gridExtra)
library(reshape2)
```

```{r caching, include=FALSE}
library("methods")
library("knitr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)

# PDF-based figures
opts_chunk$set(dev='pdf')
```


```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

opts_chunk$set(dev='pdf', dev.args=c(version="1.3"))

```

Abstract
--------

*Keywords*: forecasting, climate change, grassland, integral projection model

Introduction
------------
Population models are important tools for predicting the impacts of environmental change on species. But reconciling the scales at which population models are parameterized and the scales at which environmental changes play out remains a challenge [@Clark2010; @Freckleton2011; @Queenborough2011; @Clark2012]. The major hurdle is that most population models, at least for plant species, are built using data from small, localized plots because parameterizing traditional population models requires tracking the fates of individuals. These models are difficult to scale up from the micro to meso-scales because the fitted parameters do not fully represent the spatial variation present at scales beyond that at which the data are collected [@Sæther2007]. Thus, our ability to use population models to predict the consequences of climate change is limited when we rely on individual-level data.

Aggregate measures of individual plant performance, such as those typically collected as part of large-scale census efforts, offer an alternative to detailed demographic data for modeling populations [@Clark2004; @Freckleton2011]. Such population-level data will never match the precision of individual-level data, but it is more feasible to attain a broad coverage sample when collecting coarse-scale data. This presents a difficult trade-off: on the one hand, individual-level data leads to more reliable models; on the other hand, population-level data leads to models that will produce less precise predicitons but can be applied over greater spatial and temporal extents. If the goal of the model is to predict the consequences of climate change, then we argue that in the face of limited time and money population-level data are the best option. However, the performance of models based on population-level data have rarely been compared to that of models based on individual-level data.

To date, relatively few studies have tried to model populations based on data other than detailed individual-level data. An important exception is an effort by @Taylor2004 to model the population growth rate of an invasive species to investigate the best strategies for invasion control. They used a "density-structured" model where the state variable is a discrete density state rather than a continuous density measure. Building on this work, @Freckleton2011 showed that density-structured models compare well to continuous models in theory, and @Queenborough2011 showed the application of such methods in a study on arable weeds. In particular, @Queenborough2011 provide emprical evidence that density-structured models are capable of reproducing population dynamics, even if some precision is lost when compared to fully continuous models. Thus, population models based on coarse, population-level data show promise for producing ecological forecasts at landscape and regional scales [@Queenborough2011]. However, none of these models included environmental covariates.

Basing population models on aggregated individual-level data in a climate change context is hampered by the fact that it is individuals that respond to climate, not populations [@Clark2012]. This fact puts us in uneasy proximity to an "ecological fallacy" where one deduces inference on the individual from statistical inference on the group [@PIANTADOSI1988]. For example, individual plants may respond positively to precipitation but a negative trend is observed at the population level due to increased competition among plants as they grow larger and consume more resources. Thus, it is important to ask the question: Can aggregated data be used to detect climate signals of the same sign and magnitude as individual-level data? If not, then building population models with climate covariates on aggregated data will lead to incorrect forecasts.

Here, we test the hypothesis (idea?) that statistical and population models based on aggregated data can detect climate signals as wells as models based on individual-level data. We use a unique demographic dataset that tracks the fates of individual plants from four species to build single-species population models, since those are often used tools for ecological forecasts and climate vulnerability assessments. We first fit population models with interannual variation in vital rates explained, in part, by climate covariates. We then perturb the climate covariates to test the sensitivities of species to climate change. By doing these analyses using both individual and aggregated forms of the same data, we provide a rigorous test of our hypothesis. We find that...  

Materials and Methods
---------------------
### Study site and data
Our demographic data comes from the Fort Keogh Livestock and Range Research Laboratory in eastern Montana's northern mixed prairie near Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The dataset is freely available on Ecological Archives [@Anderson2011], and interested readers should refer to the metadata therin for a complete description. The site is about 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitaion falling from April through September. The site is grass dominated and, for the purposes of our study, we focus on the four most abundant graminoid species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE). 

From 1932 to 1945 individual plants were identified and mapped annualy in 44 1-$\text{m}^2$ quadrats using a pantograph. The quadrats were distributed in six pastures, each assigned a grazing treatment of light (1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy (0.76 ha/aum) stocking rates (two pastures per treatment). In this analysis we account for potential differences among the grazing treatments, but do not focus on grazing$\times$climate interactions. The annual maps of the quadrats were digitized and the fates of individual plants tracked and extracted using a computer program. Daily climate data, which we aggregated into climate variables of interest, are available for the duration of the data collection period (1932 - 1945) from the Miles City airport, Wiley Field, 9 km from the study site.

In this paper, we model populations based on two levels of data: individual and quadrat. The individual data is the "raw" data. For the quadrat level we data we simply sum individual areal cover for each quadrat by species. This is equivalent to a perfect census of quadrat percent cover, so we do not need to consider measurement error. Based on these two datasets we can compare population models built using individual level data and aggregated quadrat level data.

### Stastical models of vital rates
At both levels of inference (individual and quadrat), the building blocks of our population models are vital rate regressions. For individual level data we fit models for survival, growth, and recruitment of new individuals for each species. At the quadrat level we fit analagous models of extinction probability, percent cover increase/decrease, and quadrat colonization for each species. We describe the statistical models seprately since fitting the models required different approaches at the individual and quadrat levels. All models contain four climate covariate that we chose \emph{a priori}: fall through spring precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1 and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is the observation year.

We fit all models using a hierarchical Bayesian approach, which we describe in more detail below. However, for each vital rate statistical model we also define the likelihood model we use. For the likelihood models, Y is always the relevant vector of observations (e.g., whether a  genet survived [1] or not [0] from year $t$ to $t+1$). 

#### Vital rate models: individual level
We used logistic regression to model survival probability ($S$) of genet $i$ from species $j$ in quadrat group $Q$ from time $t$ to $t+1$:

\begin{align}
\text{logit}(S_{ijQ,t}) &= \gamma^{S}_{j,t} + \phi^{S}_{jQ} + \beta^{S}_{j,t}x_{ij,t} + \omega^{S}_{j}w_{ij,t} + \theta^{S}_{jk}C_{k,t} + \varepsilon^{S}_{t} \\
Y^{S}_{ijQ,t} &\sim \text{Bernoulli}(S_{ijQ,t})
\end{align}

where $x_{ij,t}$ is the log of genet size, $\gamma^{S}_{j,t}$ is a year-specific intercept, $\beta^{S}_{j,t}$ is the year-specific slope parameter for size, $\phi^{S}_{jQ}$ is the random effect of quadrat group location, and $\theta^{S}_{k}$ is the fixed parameter for the effect of the $k$th climate covariate at time $t$ ($C_{k,t}$). We include density-dependence by estimating the effect of crowding on the focal individual by other individuals of the same species. $\omega$ is the effect of crowding and $w_{t,Q}$ is the crowding experienced by the focal individual at time $t$ in quadrat group $Q$.

We modeled growth as gaussian process describing genet size at time $t+1$ as a function of size at $t$ and climate covariates:

\begin{align}
x_{ijQ,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{ij,t} + \omega^{G}_{j}w_{ij,t} + \theta^{G}_{jk}C_{k,t} \\
Y^{G}_{ijQ,t} &\sim \text{Normal}(x_{ijQ,t+1}, \sigma_{j})
\end{align}

where $x$ is log genet size and all other paramters are as described for the survival regression.

Our data allows us to track new recruits, but we cannot assign a specific parent to new genets. So, for recruitment, we work at the quadrat level and model the number of new individuals of species $j$ in quadrat $q$ recruiting at time $t+1$ as a function of quadrat "effective cover" ($A'$) in the previous year ($t$). Effective cover is a mixture of observed cover ($A$) in the focal quadrat ($q$) and the mean cover across the entire group ($\bar{A}$) of $Q$ quadrats in which $q$ is located:

\begin{equation}
A'_{jq,t} = p_{j}A_{jq,t} + (1-p_{j})\bar{A}_{jQ,t}
\end{equation}

where $p$ is a mixing fraction between 0 and 1 that is estimated within the model.

We assume the number of individuals, $Y^{R}$, recruiting at time $t+1$ follows a negative binomial distribution:

\begin{equation}
Y^{R}_{jq,t+1} \sim \text{NegBin}(\lambda_{jq,t+1},\zeta)
\end{equation}

where $\lambda$ is the mean intensity and $\zeta$ is the size parameter. We define $\lambda$ as:

\begin{equation}
\lambda_{jq,t+1} = A'_{jq,t}e^{(\gamma^{R}_{j,t} + \phi^{R}_{jQ} + \theta^{R}_{jk}C_{k,t} + \omega^{R}\sqrt{A'_{q,t}})}
\end{equation}

where $A'$ is effective cover ($\text{cm}^2$) of species $j$ in quadrat $q$ and all other terms are as in the survival and growth regressions.

#### Vital rate models: quadrat level
At the quadrat level we defined three vital rates: 

1. Probability of extirpation ($S$): the probability that, for a given species, a particular quadrat will go from non-zero cover at time $t$ to zero cover at time $t+1$.
2. Cover change ($G$): the change in percent cover from time $t$ to $t+1$ for a given species within a particular quadrat.
3. Probability of colonization ($R$): the probability that, for a give species, a particular quadrat will go from zero cover at time $t$ to non-zero cover at time $t+1$.

We retain the abbreviations $S$, $G$, and $R$ from the analagous processes at the individual level. The vital rate models at the quadrat level are all based on quadrat proportional cover. Also, at the quadrat level we do not need to explicitly include a density dependent term. Since we are modeling proportional cover, we essentially get density-dependence for "free" when proportional cover is included as a covariate. That is, density-dependence emerges because quadrats with low cover gerenally increase in cover whereas quadrats with high cover generally decrease in cover.

We modeled the probability of extirpation of species $j$ in quadrat $q$ from time $t$ to $t+1$ as:

\begin{align}
\text{logit}(S_{jq,t}) &= \gamma^{S}_{j} + \phi^{S}_{jQ} + \beta^{S}_{j}x_{jq,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{S}_{jq,t} &\sim \text{Bernoulli}(S_{jq,t})
\end{align}

where all parameters are as in Eq. 1 except that $x$ is now quadrat proportional cover. Note, however, that we do not include year random effects on the intercept $\gamma$ or the slope term $\beta$ for quadrat proportional cover. The quadrat data is inherently more sparse than the individual data from which it is aggregated, and this is especially evident when modeling rare events like extirpation and colonization. Thus, when we tried to fit random year effects, those terms did not converge.

We modeled quadrat cover change ($G$) from time $t$ to $t+1$ as:

\begin{equation}
\text{logit}(x_{jq,t+1}) = \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t}
\end{equation}

where, in this case, we do include random year effects on the intercept $\gamma$ and the slope term $\beta$. For cover change we had enough data for those terms to converge. Note that our model for quadrat cover change uses a logit transformation to link the expected cover at $t+1$ ($x_{jq,t}$) to the linear predictors. We do so because during model fitting we use a beta likelihood since the data, proportional cover, is beta distributed. The beta likelihood requires shape ($\rho$) and rate ($\eta$) parameters that can be caclulated using moment-matching:

\begin{align}
\rho_{jq,t+1} &= x_{jq,t+1}\tau_{j} \\
\eta_{jq,t+1} &= (1-x_{jq,t+1})\tau_{j}
\end{align}

with likelihood:

\begin{equation}
Y^{G}_{jq,t+1} \sim \text{Beta}(\rho_{jq,t+1}, \eta_{jq,t+1}).
\end{equation}

Finally, we modeled probability of colonization in quadrat $q$ by species $j$ from time $t$ to $t+1$ as:

\begin{align}
\text{logit}(R_{jq,t}) &= \gamma^{R}_{j} + \phi^{R}_{jQ} + \theta^{R}_{jk}C_{k,t} \\
Y^{R}_{jq,t} &\sim \text{Bernoulli}(R_{jq,t}).
\end{align}

### Model fitting
Our Bayesian approach to fitting the vital rate models required choosing appropriate priors for unknown parameters and deciding which, if any, of those prior should be hierarchical. We decided to fit models where all terms except climate covariates were fit by species, while the climate covariates were fit heierarchically where species-specific coefficients were drawn from a shared `global' coefficient distribution. We did so for two reasons: (1) the four focal species are all perrennial grasses that we expect to respond similarly to climate covariates, and (2) convergence of climate effects at the quadrat level was much easier to achieve when we modeled these terms hierarchically, allowing them to "share" statistical strenght via partial pooling [@Gelman2007]. So, climate effects were modeled as:

\begin{align}
\theta_{jk} &\sim \text{Normal}(\bar{\theta_{k}}, \sigma_{k})
\end{align}

where $\bar{\theta_{k}}$ is the interspecific effect of the $k$th climate covariate.

We used uninformative priors for all unknown parameters, specifically:

\begin{align}
\boldsymbol{\gamma, \beta, \bar{\theta}} &\sim \text{Normal}(0, 1e^{-6}) \\
\boldsymbol{\phi} &\sim \text{Normal}(0, \boldsymbol{\sigma_{\phi}}) \\
\boldsymbol{\sigma_{\phi}} &\sim e^{(\text{Gamma}(2, 0.5))} \\
\boldsymbol{\sigma_{\theta}, \sigma_{\gamma}, \sigma_{\beta}, \tau, \zeta} &\sim \text{Gamma}(0.001, 0.001)
\end{align}

All of our analyses (model fitting and simulating) were conducted in R [@Team2013]. We used the MCMC sampler in JAGS [@Plummer2003] to estimate the posterior distributions of model parameters and the package 'r2jags' [@Su2012] to connect R to JAGS. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 50,000 iterations, after discarding an initial 50,000 iterations. We assessed convergence visually and using the @Gelman1992 diagnostic in the R package `coda' [@Plummer2006]. Scale reduction factors for all parameters were less than 1.02, indicating convergence. For the purposes of introducing stochasticity in our population models, we saved the final 1,000 iterations from each chain for all parameters to be used as randomly drawn values during population simulation.

We assessed the statistical importance of the climate covariates included the final vital rate regressions by comparing the residual deviance of models with climate covariates and temporal random effects, climate covariates only, and temporal random effects only. When a model includes climate covariates, this comparison shows the relative contribution of the climate covariates in explaining the total interannual variability [@Adler2012]. 

### Population models
With the posterior distribution of the vital rate statistical models in hand, it is straightforward to simulate the population models. We used an Integral Projection Model (IPM) to model populations based on individual level data and an quadrat based version of an individually-based model (Quadrat-Based Model, QBM) to model populations based on quadrat level data. Both models take the general form:

\begin{equation}
N_{t+1} = S \times G + R.
\end{equation}

So, at each time step in a simulation, we use the survival regression to determine if each genet lives or not (if each quadrat remains occupied or not), the growth regression to determine size changes of surviving individuals (cover change of occupied quadrats), and the recruitment regression to determine the number of new recruits (if a quadrat is colonized or not). We first use one-step-ahead forecasts to assess each model's ability to reproduce observed cover changes. Then we use the models to analyze the effect of potential climate changes on population size over long time scales.

We used random draws from the final 1,000 iterations from each of 3 MCMC chains to introduce stochasticity into our population models. At each time step, we first randomly selected climate covariates from one of the 14 observed years and also randomly drew one set of random year effects for each vital rate regression. Then, we drew the full parameter set (specific random year effects, climate effects, fixed effects) from a randomly selected MCMC iteration. We selected random year effects and the climate year independently. 

Results
-------
### Climate effects on vital rates
At the individual level we fit vital rate models that included climate covariates and random year effects (full model), just climate covariates (climate model), and no climate or random year effects (constant model). We used the metric from @Adler2012 to quantify the contribution of climate covariates to explaining interannual variability in vital rates based on comparisons of model deviance: (Climate model - Constant Model)/(Full model - Constant model). At the quadrat level, we could only fit random year effects for the growth regression. This is, in fact, one of the limitation of census-based data: local extirpation and colonization events are likely to be rare, making it difficult to fit complex statistical models. So, in this section we focus on growth only.

Climate covariates improved the growth regressions at both levels, with a slghty greater contribution at the individual level (Figure 1). Parameter estimates were generally consistent in terms of sign and magnitude between the individual level and the quadrat level for each species and climate covariate (Figure 2). There is no clear pattern suggesting that parameters estimated using individual level data are more certain, as evidenced by credible interval widths, than estimates based on quadrat data.

### Climate effects on population dynamics

Discussion
----------
We sought to test the assumption that the sensitivities of plant populations to climate variables can be detected equally well using either individual level data or population level data. This is an important question to answer because population models are key tools for predicting the consequences of global climate change. However, they can be of limited use when built on data from a small subset of a population in space or time. If population level data (i.e., some aggregated form of individual level data) can be used to detect climate effects on population dynamics, then we would have a cheaper and easier option for data collection over relatively large temporal and spatial extents [_e.g._ @Freckleton2011].

1. Climate detections
2. Limitations:
  * Sample size at quadrat level
  * are these species inherently sensitive to climate (compare to Chu and Adler in review -- say why we chose Montana site -- most sensitive to climate variables)
  * we know recruitment is important, and this is something that may be easily missed at quad level
3. Implications

\pagebreak{}

Figures
-------
```{r figure_1, dependson="plot-options", fig.cap="Boxplots of model residuals for one-step-ahead forecasts at each observation year. Each one-step forecast was simulated `r nSims` times. Note that the y-axes vary across panels.", fig.width=5, fig.height=6, cache=FALSE, results='hide'}
resD <- readRDS("../montana/quadBM/simulations/quadBM_oneStep_Residuals.rds")
nSims <- length(unique(resD$sim))
diffD <- ddply(resD, c("year", "species"), summarise,
               avgDiff = mean(yearDiff))

ggplot(resD)+
  geom_hline(aes(yintercept=0))+
#   geom_line(data=diffD, aes(x=year, y=avgDiff, group=species), color="purple", alpha=0.5)+
#   stat_smooth(data=diffD, aes(x=year, y=avgDiff, group=species), color="lightblue", se = FALSE, size=1)+
  geom_boxplot(aes(x=year, y=covChangeResiduals), outlier.shape = 1, width=0.3, outlier.size = 0.5)+
  facet_grid(species~., scales = "free_y")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r figure_2, dependson="plot-options", fig.cap="Simulated changes species' median cover caused by a 1% increase in observed precipitation or temperature. We used the median of simulated results because the average was highly influence by some extreme outliers.", fig.width=4.5, fig.height=4, cache=FALSE, results='hide'}
climBase <- readRDS("../montana/quadBM/simulations/climateChangeBase.rds")
climPpt <- readRDS("../montana/quadBM/simulations/climateChangePrecip.rds")
climTemp <- readRDS("../montana/quadBM/simulations/climateChangeTemperature.rds")
climBoth <- readRDS("../montana/quadBM/simulations/climateChangePptTmp.rds")
climBase$sim <- rep("Observed", nrow(climBase)) 
climPpt$sim <- rep("Precipitation", nrow(climBase)) 
climTemp$sim <- rep("Temperature", nrow(climBase)) 
climBoth$sim <- rep("Both", nrow(climBase)) 
climSimsQBM <- rbind(climBase, climPpt, climTemp, climBoth)
# ggplot(climSimsQBM)+
#   geom_boxplot(aes(x=species, y=cover*100, fill=sim))+
#   scale_y_continuous(limits=c(0,10))
climSimsQBMavg <- ddply(climSimsQBM, .(sim, species), summarise,
                        avgCover = median(cover))
climSimQBMcast <- dcast(climSimsQBMavg, formula = species~sim)
climSimQBMcast$diffP <- with(climSimQBMcast, (Precipitation*100)-(Observed*100)) 
climSimQBMcast$diffT <- with(climSimQBMcast, (Temperature*100)-(Observed*100))
climSimQBMcast$diffB <- with(climSimQBMcast, (Both*100)-(Observed*100))
climSimQBMplot <- melt(climSimQBMcast)
ggplot(data=subset(climSimQBMplot, variable=="diffP"|variable=="diffT"|variable=="diffB"))+
  geom_bar(aes(x=species, y=value, fill=variable), stat="identity", position="dodge", color="white")+
  scale_fill_manual(name="", labels=c("1% Precipitation Increase", "1% Temperature Increase", "1% Temp. & Ppt. Increase"), values=c("grey25", "grey50", "grey75"))+
  theme_bw()+
  theme(legend.position=c(0.6, 0.8))+
  ylab("Cover Change (%)")+
  xlab("Species")
```

\pagebreak{}

References
----------
